\vspec{} and Friends: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.

### 2024-07-30 src/tex/ms.tex
Rewrite facula area section

\subsection{Faculae \label{subsec:faculae}}
Faculae are magnetically-generated regions of the solar surface that usually appear as bright points near the limb; we employ the ``hot wall''
model \citep{spruit1976} where faculae are described as three-dimensional pores in the stellar surface with a hot, bright wall and a cool, dark floor, as shown in Figure \ref{fig:fac_struct}. Their three-dimensional structure causes faculaes' observational properties to change depending on their angle from disk-center. Close to the limb,
the hot wall is visible to the observer, and faculae appear as bright points; near the center, however, the cool floor is exposed and faculae appear dark. To consider this effect in the faculae lightcurve, we compute the fraction of each facula's projected area -- the area on the disk it would occupy as seen by the observer -- that is occupied by the hot wall versus the cool floor. This is done via numerical integral along the radius of the spot.

We calculate each portion using geometrical relationships between the spot dimensions and the viewing angle:
\begin{equation}
    f_{wall} = \frac{\int _{-R}^{R} Z_{\rm eff} dr}{\int _{-R}^{R} Z_{\rm eff} dr + \int _{-R}^{R} R_{\rm eff} dr}
\end{equation}
where $Z_{\rm eff}$ is the apparent height of the wall
\begin{equation}
    Z_{\rm eff} = \left\{
    \begin{array}{lr}
        Z_w \sin{\alpha}, & \text{if } \alpha \leq \alpha_{\rm crit} \\
        2\sqrt{R^2 - r^2}\cos{\alpha}, & \text{if } \alpha > \alpha_{\rm crit}
    \end{array}
    \right\}
\end{equation}
and $R_{\rm eff}$ is the apparent width of the floor
\begin{equation}
    R_{\rm eff} = \left\{
    \begin{array}{lr}
        2\sqrt{R^2 - r^2} - Z_w\sin{\alpha}, & \text{if } \alpha \leq \alpha_{\rm crit} \\
        0, & \text{if } \alpha > \alpha_{\rm crit}
    \end{array}
    \right\}
\end{equation}
for facula radius $R$, depth $Z_w$, and angle from disk-center $\alpha$. $r$ in this numerical scheme is defined as the distance from the center of the facula along the radial line connecting the facula center to the disk center. $\alpha_{\rm crit}$ is the value of alpha at which the floor is no longer visible and is defined to be $\arctan{\frac{2\sqrt{R^2-r^2}}{Z_w}}$.

According to studies of solar faculae \citep{topka1997}, faculae temperatures (of both the floor and wall) are dependent on the facula radius, while depth appears to be constant. They also find that the smallest faculae have no visible floor, and that even at disk center they appear as bright points. We parameterize the floor temperature to be
\begin{equation}
    \Delta T_{\rm eff, floor} = \left\{
    \begin{array}{lr}
    \text{Not visible}, & \text{if } r<r_{\rm min} \\
    m_{\rm floor}(r-r_{\rm min}) + \Delta T_{\rm eff, floor, 0}, & \text{if } r \ge r_{\rm min}
    \end{array}
    \right\}
\end{equation}
\TJ{Change variable names, make diagram}
where $r$ is the radius of the facula, $r_{\rm min}$ is the minimum radius where the floor is visible, $\Delta  T_{\rm eff, floor, 0}$ is the difference between the floor and photosphere \teff at $r_{\rm min}$, and $m_{\rm floor}$ is the slope of the relationship with units of [temperature] [length]$^{-1}$. Similarly, the wall temperature is parameterized as
\begin{equation}
    \Delta T_{\rm eff,wall} = m_{\rm wall}r + \Delta T_{\rm eff,wall,0}
\end{equation}
where $\Delta T_{\rm eff,wall,0}$ is the temperature of a zero-radius facula and $m_{\rm wall}$ is the slope of the radius-temperature relationship with units of [temperature] [length]$^{-1}$. These relationships can be defined by the user, for example setting $m=0$ for constant temperatures.

@@

\begin{figure}
    \centering
    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_model.png}}
    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_depth.pdf}}
    \script{facula_depth.py}
    \caption{
        ``Hot Wall'' model of faculae. Faculae structure causes their contrast to be dependent on their distance from the center of the disk. {\bf Top}: Depiction of a facula on the limb of a star. The hot wall is exposed to the observer causing the pore to appear bright. At disk center, the cool floor is most visible. {\bf Bottom}: The effects of depression depth and viewing angle on facula brightness. The 3D structure of faculae is most apparent when radius $\sim$ depth. A toy flux model was used to demonstrate the shape of these curves, but in practice their magnitudes depend on stellar spectral models.
        }
    \label{fig:fac_struct}
\end{figure}

>>>>

\subsection{Faculae \label{subsec:faculae}}
Faculae are magnetically-generated regions of the solar surface that usually appear as bright points near the limb; we employ the ``hot wall''
model \citep{spruit1976} where faculae are described as three-dimensional pores in the stellar surface with a hot, bright wall and a cool, dark floor, as shown in Figure \ref{fig:fac_struct}. Their three-dimensional structure causes faculaes' observational properties to change depending on their angle from disk-center. Close to the limb,
the hot wall is visible to the observer, and faculae appear as bright points; near the center, however, the cool floor is exposed and faculae appear dark. To consider this effect in the faculae lightcurve, we compute the fraction of each facula's projected area -- the area on the disk it would occupy as seen by the observer -- that is occupied by the hot wall versus the cool floor. Mathematically, this is no different from finding the area of the intersection of two overlapping ellipses:

As shown in Figure \ref{fig:fac_struct}, peering through the top of the pore the observer can see a portion of the hot wall and a portion of the cool floor. Consider an ellipse centered at the origin with semimajor axis $R$ and semiminor axis $\mu R$, oriented so that the semimajor axis points along the $y$-axis. Now consider a second, identical ellipse, displaced along the $x$-axis by some amount $0<d<\mu R$. The intersection of these two ellipses is the portion of the cool floor that is visible, and the area of the second ellipse outside the intersection is the portion of the hot wall that the observer sees. We are not interested in the absolute areas of each of these regions, only how they relate to the total ellipse area $\pi \mu R^2$, and so we can generalize to the case that $\mu=1$ (i.e. circles). This problem is solved for arbitrary circles by \citet{weisstein2004}. The details of their derivation are not relevant to this text, but we use their result to create a function $f(d,R)$ that computes the fraction of a circle with radius $R$ and center $(d,0)$ that is inside the unit circle. For a facula with radius $r$, wall height $h$, and angle from disk center $theta$, we find that the effective area of the cool floor is:
\begin{equation}
    \frac{A_{\rm floor}}{A_{\rm total}} = f\left(\frac{h}{r}\tan{\theta}, 1 \right)
\end{equation}
and for the hot wall it is:
\begin{equation}
    \frac{A_{\rm wall}}{A_{\rm total}} = 1-  f\left(\frac{h}{r}\tan{\theta}, 1 \right)
\end{equation}

According to studies of solar faculae \citep{topka1997}, faculae temperatures (of both the floor and wall) are dependent on the facula radius, while depth appears to be constant. They also find that the smallest faculae have no visible floor, and that even at disk center they appear as bright points. We parameterize the floor temperature to be
\begin{equation}
    \Delta T_{\rm eff, floor} = \left\{
    \begin{array}{lr}
    \text{Not visible}, & \text{if } r<r_{\rm min} \\
    m_{\rm floor}(r-r_{\rm min}) + \Delta T_{\rm eff, floor, 0}, & \text{if } r \ge r_{\rm min}
    \end{array}
    \right\}
\end{equation}
where $r$ is the radius of the facula, $r_{\rm min}$ is the minimum radius where the floor is visible, $\Delta  T_{\rm eff, floor, 0}$ is the difference between the floor and photosphere \teff at $r_{\rm min}$, and $m_{\rm floor}$ is the slope of the relationship with units of [temperature] [length]$^{-1}$. Similarly, the wall temperature is parameterized as
\begin{equation}
    \Delta T_{\rm eff,wall} = m_{\rm wall}r + \Delta T_{\rm eff,wall,0}
\end{equation}
where $\Delta T_{\rm eff,wall,0}$ is the temperature of a zero-radius facula and $m_{\rm wall}$ is the slope of the radius-temperature relationship with units of [temperature] [length]$^{-1}$. These relationships can be defined by the user, for example setting $m=0$ for constant temperatures.

@@

\begin{figure}
    \centering
    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_diagram.pdf}}
    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_depth.pdf}}
    \script{facula_depth.py}
    \caption{
        ``Hot Wall'' model of faculae. Faculae structure causes their contrast to be dependent on their distance from the center of the disk. {\bf Top}: Depiction of a facula on the limb of a star. The hot wall is exposed to the observer causing the pore to appear bright. At disk center, the cool floor is most visible. {\bf Bottom}: The effects of depression depth and viewing angle on facula brightness. The 3D structure of faculae is most apparent when radius $\sim$ depth. A toy flux model was used to demonstrate the shape of these curves, but in practice their magnitudes depend on stellar spectral models.
        }
    \label{fig:fac_struct}
\end{figure}
